\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-28 Draft_Form/tex/Results.tex
update 28_02_2018
update 28_02_2018

On both small area levels, the design-unbiased estimators \psmall{} and \extpsynth{} led to a substantial reduction in the estimation error compared to the SRS estimator (Fig. \ref{fig:disterrors}). On the FA level, the SRS estimator yielded an estimation error of 6.7\% on average compared to 5.2\% and 4.8\% under the \extpsynth{} and \psmall{} respectively (Table \ref{tab:estres}). The cumulative error distribution (Fig. \ref{fig:disterrors}, left) reveals that under the SRS estimator, errors less than 5\% were achieved for 17\% of the FA units (8 of 45). This proportion could be increased to 62\% (28 FA units) and 73\% (33 FA units) by application of the \psmall{} and \extpsynth{} estimator. 95\% of all estimates exhibited errors less than 9.5\% under the SRS estimator and less than 6.6\% when using \psmall{} or \extpsynth{}. Estimation errors higher than 10\% only appeared twice for each of the three estimators.\par
Although the estimation errors were substantially larger overall on the FR level compared to the FA level due to smaller sample sizes, the error reduction from SRS by \psmall{} and \extpsynth{} were even more pronounced (Fig. \ref{fig:disterrors}, right). The average error under the SRS estimator was 18.3\%, while it was 11.3\% and 12.2\% under \psmall{} and \extpsynth{} (Table \ref{tab:estres}). Errors smaller than 10\% were achieved for 15\% of the FR units by the SRS estimator, and for 46\% by the \psmall{} and \psynth{} estimator. 95\% of the 321 FR units where \psmall{} and \extpsynth{} could be applied exhibited errors less than 20\%. In comparion, the SRS estimates resulted in errors less than 36.6\% for 95\% of the 388 FR units.

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/error_distr_foa_fu.png}}
	\caption{Cumulative distribution of estimation errors under SRS, \psmall{}, \extpsynth{} and the \psynth{} estimator. \textit{Left}: Results for the 45 FA units. \textit{Right}: Results for the 388 (SRS), 321 (\psmall{} / \extpsynth{}) and 403 (\psynth{}) FR units.}
	\label{fig:disterrors}
\end{figure}

On both small area levels, the \psynth{} estimator resulted in much smaller estimation errors compared to \psmall{} and \extpsynth{}. This was as expected, since the \psynth{} variance estimate does not take the residual variation in each small area unit into account (section \ref{sec:psynth}). Compared to the asymptotically design-unbiased estimators \psmall{} and \extpsynth{}, the estimation errors produced by \psynth{} thus seem to be too optimistic. One should also recall that the estimates of the \psynth{} estimator are potentially design-biased.

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in section \ref{sec:extpsynth}, one possible explanation for such differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%).

However, the affected FR units (depicted in red diamonds, fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates.\par

@@

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/psmall_vs_extpsynth_fu.png}}
	\caption{\textit{Left}: Comparison of the g-weight variance between the PSMALL and the EXTPSYNTH estimator for the 321 FR units.
		\textit{Right}: Difference in g-weight variance between the PSMALL and the EXTPSYNTH estimator in dependence of the terrestrial data ($n2G$) in the FR unit.}
	\label{fig:compvar}
\end{figure}

@@

The variance reduction relative to SRS for \psmall{} and \extpsynth{} are described in Figure \ref{fig:gain} and Table \ref{tab:gain}. A direct comparison of the variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{}, \extpsynth{}) led to a variance reduction compared to SRS in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the variance by up to 54.1\%. The reduction in variance can also be expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR level, the reduction in variance and the relative efficiencies reached even higher values (Table \ref{tab:gain} and Fig. \ref{fig:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the generally smaller variances of the \extpsynth{} estimator (section \ref{sec:comp}).

@@

Cases also occurred on the FR level where one or both two-phase estimators produced larger variance values than under the SRS estimator. This happened in 19\% of the FR units under the \extpsynth{}, and in 24\% of the FR units under the \psmall{} estimator. One possible reason for this was supposed to be a large residual variance due to a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the \psmall{} estimator (eq. \ref{eq:psmallvar}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residual (term 3). In general, the residual term is expected to make the largest contribution to the overall variance since it's sample size is based on $n_{2,G}$ whereas the auxiliary term and the coefficient term are based on larger sample sizes, i.e. $n_{1,G}$ and $n_2$ respectively. Figure \ref{fig:fail} illustrates the share of the overall variance by the residual term of the \psmall{} estimator scaled by the overall percentage reduction or increase of the variance compared to SRS for various small area sample sizes $n_{2,G}$.\par

The residual term generally constitutes the dominating part of the \psmall{} variance (around 84\% on average). Although high residual term dominance does not necessary indicate that the \psmall{} variance will be disproportionately large, as apparent from Figure \ref{fig:fail} (\textit{right}), the vast majority of the small areas where the \psmall{} variance was larger than the SRS variance had residual terms contributing over 75\% to the overall \psmall{} variance. Furthermore, the magnitudes of the worst cases tended to occur in lower sample sizes. For example, of the FR units that saw variance increases where $n_{2,G}=4$, the average increase was 272\%, compared to 62\% for FR units with $n_{2,G} > 4$ (Fig. \ref{fig:fail}, \textit{left}).  In comparison, the magnitude of the variance decreases were far more homogeneous than for the variance increases regardless of terrestrial sample size. Since $n_{2,G}$ is the same for \psmall{} and SRS, this implies that the sum of square residuals for the model are likely larger than the sum of square local densities for the clusters in $s_{2,G}$ indicating the presence of outliers with large residuals in the problematic small areas. This situation is likely to arise when there was forest loss after the ALS scanning but before the terrestrial survey year. % [check that]

@@

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/eval_2phase_fail_11.png}}
	\caption{Increase or reduction of the SRS variance by application of the \psmall{} estimator for all 321 FR units in dependence of a) the terrestrial sample size in the FR units (\textit{x-axis}) and b) the residual term proportion of the \psmall{} g-weight variance (\textit{y-axis}).}
	\label{fig:fail}
\end{figure}

>>>>

On both small area levels, the design-unbiased estimators \psmall{} and \extpsynth{} led to a substantial reduction in the estimation error compared to the SRS estimator (Fig. \ref{fig:disterrors}). On the FA level, the SRS estimator yielded an estimation error of 6.7\% on average compared to 5.2\% and 4.8\% under \extpsynth{} and \psmall{} respectively (Table \ref{tab:estres}). The cumulative error distribution (Fig. \ref{fig:disterrors}, left) reveals that under the SRS estimator, errors less than 5\% were achieved for 17\% of the FA units (8 of 45). This proportion could be increased to 62\% (28 FA units) and 73\% (33 FA units) by application of the \psmall{} and \extpsynth{} estimator. 95\% of all estimates exhibited errors less than 9.5\% under the SRS estimator and less than 6.6\% when using \psmall{} or \extpsynth{}. Estimation errors higher than 10\% only appeared twice for each of the three estimators.\par
Although the estimation errors were substantially larger overall on the FR level compared to the FA level due to smaller sample sizes, the error reduction from SRS by \psmall{} and \extpsynth{} were even more pronounced (Fig. \ref{fig:disterrors}, right). The average error under the SRS estimator was 18.3\%, while it was 11.3\% and 12.2\% under \psmall{} and \extpsynth{} (Table \ref{tab:estres}). Errors smaller than 10\% were achieved for 15\% of the FR units by the SRS estimator, and for 46\% by the \psmall{} and \psynth{} estimator. 95\% of the 321 FR units where \psmall{} and \extpsynth{} could be applied exhibited errors less than 20\%. In comparion, the SRS estimates resulted in errors less than 36.6\% for 95\% of the 388 FR units.

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/error_distr_foa_fu.png}}
	\caption{Cumulative distribution of estimation errors under SRS, \psmall{}, \extpsynth{} and the \psynth{} estimator. \textit{Left}: Results for the 45 FA units. \textit{Right}: Results for the 388 (SRS), 321 (\psmall{}, \extpsynth{}) and 403 (\psynth{}) FR units.}
	\label{fig:disterrors}
\end{figure}

On both small area levels, the \psynth{} estimator resulted in much smaller estimation errors compared to \psmall{} and \extpsynth{}. This was as expected, since the \psynth{} variance estimate does not take the residual variation in each small area unit into account (Section \ref{sec:psynth}). Compared to the asymptotically design-unbiased estimators \psmall{} and \extpsynth{}, the estimation errors produced by \psynth{} thus seem to be too optimistic. One should also recall that the estimates of the \psynth{} estimator are potentially design-biased.

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (Fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in Section \ref{sec:extpsynth}, one possible explanation for differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%). We compared the variances of \psmall{} and \extpsynth{} for all small areas that did not have the violations using a Wilcoxon Rank-Sum Test \citep{wilcoxon1970}. This test was also performed for groups $n_{2,G} \leq 6$, $n_{2,G} > 6$ and $n_{2,G} > 10$. The distribution of variances from \extpsynth{} were found to be significantly lower than that of \psmall{} (p-values of 1.2*10$^{-13}$, 3.4*10$^{-8}$) except for the group of $n_{2,G} > 10$ (p-value of 0.125). The latter was expected since the variances of both estimators are asymptotically equivalent under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. This was also confirmed by a comparison of the absolute differences in the variances (Fig. \ref{fig:compvar}, right) which decreased with increasing terrestrial sample size. The Wilcoxon Rank-Sum Test was also used to compare the \extpsynth{} variances of those small areas with violation (depicted in red diamonds, Fig. \ref{fig:compvar}) to the \psmall{} variances without violations. Overall as well as for $n_{2,G} \leq 6$ there was a significant difference, but not for $n_{2,G} > 6$. This provided some evidence that the violations created a statistically significant influence on the \extpsynth{} variance that makes it appear to be slightly over-optimistic. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates.\par

%However, the affected FR units (depicted in red diamonds, Fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (Fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates.\par

@@

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/psmall_vs_extpsynth_fu.png}}
	\caption{\textit{Left}: Comparison of the g-weight variance between the PSMALL and the EXTPSYNTH estimator for the 321 FR units.
		\textit{Right}: Difference in g-weight variance between the PSMALL and the EXTPSYNTH estimator in dependence of the terrestrial data ($n_{2,G}$) in the FR unit.}
	\label{fig:compvar}
\end{figure}

@@

The variance reduction relative to SRS for \psmall{} and \extpsynth{} are described in Figure \ref{fig:gain} and Table \ref{tab:gain}. A direct comparison of the variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{}, \extpsynth{}) led to a variance reduction compared to SRS in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the variance by up to 54.1\%. The reduction in variance can also be expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR level, the reduction in variance even reached values of 90\% and relative efficiencies of 30 (Table \ref{tab:gain} and Fig. \ref{fig:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the generally smaller variances of the \extpsynth{} estimator (Section \ref{sec:comp}).

@@

Cases also occurred on the FR level where one or both two-phase estimators produced larger variance values than under the SRS estimator. This happened in 19\% of the FR units under the \extpsynth{}, and in 24\% of the FR units under the \psmall{} estimator. One possible reason for this was supposed to be a large residual variance due to a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the \psmall{} estimator (eq. \ref{eq:psmallvar}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residuals (term 3). In general, the residual term is expected to make the largest contribution to the overall variance since it's sample size is based on $n_{2,G}$ whereas the auxiliary term and the coefficient term are based on larger sample sizes, i.e. $n_{1,G}$ and $n_2$ respectively. Figure \ref{fig:fail} illustrates the share of the overall variance by the residual term of the \psmall{} estimator scaled by the overall percentage reduction or increase of the variance compared to SRS for various small area sample sizes $n_{2,G}$. The residual term generally constitutes the dominating part of the \psmall{} variance (around 84\% on average). Although high residual term dominance does not necessary indicate that the \psmall{} variance will be disproportionately large, as apparent from Figure \ref{fig:fail} (\textit{right}), the vast majority of the small areas where the \psmall{} variance was larger than the SRS variance had residual terms contributing over 75\% to the overall \psmall{} variance. Furthermore, the magnitudes of the worst cases tended to occur in lower sample sizes. For example, of the FR units that saw variance increases where $n_{2,G}=4$, the average increase was 272\%, compared to 62\% for FR units with $n_{2,G} > 4$ (Fig. \ref{fig:fail}, \textit{left}).  In comparison, the magnitude of the variance decreases were far more homogeneous than for the variance increases regardless of terrestrial sample size. Since $n_{2,G}$ is the same for \psmall{} and SRS, this implies that the sum of square residuals for the model are likely larger than the sum of square local densities for the clusters in $s_{2,G}$ indicating the presence of outliers with large residuals in the problematic small areas. This situation is likely to arise when there was forest loss after the ALS scanning but before the terrestrial survey year. % [check that]

@@

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/eval_2phase_fail_11.png}}
	\caption{Share of the overall variance by the residual term of the \psmall{} estimator for various small area sample sizes. Points are scaled by the overall percentage reduction/increase of the variance compared to SRS.}
	\label{fig:fail}
\end{figure}
